Text embedding models yield high resolution insights into conceptual knowledge from short multiple choice quizzes

### 2023-02-01 paper/main.tex
backing up

We hypothesized that a topic model trained on transcripts of the two lectures,
should also capture the conceptual knowledge probed by each quiz question. If
indeed the topic model could capture information about the deeper conceptual
content of the lectures (i.e., beyond surface-level details such as particular
word choices) then we should be able to recover a correspondence between each
lecture and questions \textit{about} each lecture. Importantly, such a
correspondence could not solely arise from superficial text matching between
lecture transcripts and questions, since the lectures and questions used
different words. Simply comparing the average topic weights from each lecture
and question sets (averaging across time and questions, respectively) reveals a
striking correspondence (Fig.~\topicWeights). Specifically, the average topic
weights from lecture 1 are strongly correlated with the average topic weights
from lecture 1 questions ($r(13) = XX, p = XX$, 95\% confidence interval
(CI)~$= XX$), and the average topic weights from lecture 2 are strongly
correlated with the average topic weights from lecture 2 questions ($r(13) =
XX, p = XX, CI = XX$). At the same time, the average topics from two lectures
are \textit{negatively} correlated ($r(13) = XX, p = XX, CI = XX$). The full
set of pairwise comparisons between topic vectors for the lectures and each
question set is reported in Figure~\topicWeights.

>>>>

We hypothesized that a topic model trained on transcripts of the two lectures,
should also capture the conceptual knowledge probed by each quiz question. If
indeed the topic model could capture information about the deeper conceptual
content of the lectures (i.e., beyond surface-level details such as particular
word choices) then we should be able to recover a correspondence between each
lecture and questions \textit{about} each lecture. Importantly, such a
correspondence could not solely arise from superficial text matching between
lecture transcripts and questions, since the lectures and questions used
different words. Simply comparing the average topic weights from each lecture
and question sets (averaging across time and questions, respectively) reveals a
striking correspondence (Fig.~\topicWeights). Specifically, the average topic
weights from lecture 1 are strongly correlated with the average topic weights
from lecture 1 questions ($r(13) = 0.809, p < 0.001$, 95\% confidence interval
(CI)~$= [0.633, 0.962]$), and the average topic weights from lecture 2 are
strongly correlated with the average topic weights from lecture 2 questions
($r(13) = 0.728, p = 0.002, CI = [0.456, 0.920]$). At the same time, the
average topics from two lectures are \textit{negatively} correlated ($r(13) =
-0.634, p = 0.011, CI = [-0.924, -0.237]$), indicating that the topic model
also exhibits some degree of specificity. The full set of pairwise comparisons
between topic vectors for the lectures and each question set is reported in
Figure~\topicWeights.
